---
title: "Day Interactive"
format: html
editor: visual
---

## Data wrangling refresher

```{r}
library(tidyverse)
library(palmerpenguins)
```

1.  starting from penguins, write a single piped sequence in which you include penguins at Dream and Biscoe Islands, remove the year and sex variables, add a new column body mass kg, with penguin mass converted from grams to kg, and rename the island variable to location

```{r}
penguins_sub1 <- penguins %>% 
  filter(island %in% c("Dream", "Biscoe")) %>% 
  select(-year, -sex) %>% 
  mutate(body_mass_kg = body_mass_g / 1000) %>% 
  rename(location = island)
```

starting from penguins, write a single piped sequence to:

-   limit to adelie penguins

-   remove any observations where flipper_length_mm is NA (!is.na() or drop_na()

-   group the data by sex

-   create a summary table that contains, mean, standard deviation, and sample size of flipper lengths for male and female Adelie penguins

```{r}
penguins_sub2 <- penguins %>% 
  filter(species == "Adelie") %>% 
  drop_na(flipper_length_mm, sex) %>% 
  group_by(sex) %>% 
  summarize(mean_length = mean(flipper_length_mm),
            sd_length = sd(flipper_length_mm),
            sample_size = n())

# Or, using !is.na(), which can only use one input
penguins_sub3 <- penguins %>% 
  filter(species == "Adelie", 
         !is.na(flipper_length_mm),
         !is.na(sex)) %>% 
  group_by(sex) %>% 
  summarize(mean_length = mean(flipper_length_mm),
            sd_length = sd(flipper_length_mm),
            sample_size = n())
```

Using count()

```{r}
# count() wraps group_by and summarize and n to get counts of observations. If you want summary statistics, you still want to use all 3 together
penguins %>% 
  count(species, island, sex)
```
